Nested Stochastic Block Models Applied to the Analysis of Single Cell Data

### 2021-07-19 manuscript.tex
Update manuscript.tex

\textcolor{red}{Only by reducing resolution to $\gamma \leq 0.5$ we were able to obtain a single partition by modularity (Fig. \ref{Modularity_Random_Res}). Of course,} this experiment is a deliberate extreme case. The quality of grouping proposed by a standard approach can be disputed in many ways, and the UMAP embedding indeed reflects the absence of any information. Nevertheless, real-world data may include an unknown amount of random noise. Hence, It is important to identify cell groups that are not artefacts arising from processing and that do reflect the information contained in the dataset. \textcolor{red}{To understand the impact of random noise on structured data, we considered the same PBMC dataset and added white noise to the normalised counts at increasing levels of $\sigma$, ensuring that the noise level is modeled after the feature-wise distribution of detected genes. We then compared partitions to the original annotation by Adjusted Rand Index ($ARI$), we found that \texttt{schist} is more robust to perturbations and that, again, optimising the modularity results in overestimation of the number of communities at high noise levels (Table \ref{Table_Noise}. Of note, the concordance with original annotations drops at $\sigma \geq 1.5$ for both the approaches.}

>>>>

\textcolor{red}{Only by reducing resolution to $\gamma \leq 0.5$ we were able to obtain a single partition by modularity (Fig. \ref{Modularity_Random_Res}). Of course,} this experiment is a deliberate extreme case. The quality of grouping proposed by a standard approach can be disputed in many ways, and the UMAP embedding indeed reflects the absence of any information. Nevertheless, real-world data may include an unknown amount of random noise. Hence, it is important to identify cell groups that are not artefacts arising from processing and that do reflect the information contained in the dataset. \textcolor{red}{To understand the impact of random noise on structured data, we considered the same PBMC dataset and added white noise to the normalised counts at increasing levels of $\sigma$, ensuring that the noise level is modeled after the feature-wise distribution of detected genes. We then compared partitions to the original annotation by Adjusted Rand Index ($ARI$), we found that \texttt{schist} is more robust to perturbations and that, again, optimising the modularity results in overestimation of the number of communities at high noise levels (Table \ref{Table_Noise}. Of note, the concordance with original annotations drops at $\sigma \geq 1.5$ for both the approaches.}
